
---
title: "Assignment1"
output: html_document
---
# Abstract 

This document attempts to quantify "how well they do" weight lifting exercise.  The dataset for this document is based on [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har).

This report outlines the steps for the prediction model development. 

# Loading and Preprocessing of the data

```{r}
set.seed(2334)
library(caret)
library(randomForest)
library(doParallel)
cl <- makeCluster(detectCores())
registerDoParallel(cl)
```

Loading of the dataset.  Also in this loading step, the following "NA", "#DIV/0!" is assumed to be NA.
```{r}
data <- read.csv("./pml-training.csv", na.strings=c("NA", "#DIV/0!") )
testing <- read.csv("./pml-testing.csv", na.strings=c("NA"))
```

For the purposes of this model development.  The following columns ("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window") are also removed.  These columns would introduce a lot more noise.
```{r}
noisycols <- c("X", "user_name", "raw_timestamp1_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")
data <- data[,-which(names(data) %in% noisycols)]
testing <- testing[,-which(names(testing) %in% noisycols)]
```

Also removed are columns which have NA as its values.
```{r}
nacols <- apply(data, 2, function(x) {
     sum(is.na(x))
})
data <- data[,which(nacols == 0)]
testing <- testing[,which(nacols == 0)]
```


# Partitioning of the data

A 70% data split is used to partition the data between training set and test set

```{r}
inTrain <- createDataPartition(y=data$classe, p=0.7, list=FALSE)
trainset <- data[inTrain,]
testset <- data[-inTrain,]
```

# Prediction Model Development

For this exercise, random forest model is chosen, given the higher accuracy achieved.  For efficiency (faster run-time), the tree has been limited to 50. 

```{r}
data.rf <- randomForest(classe ~., data=trainset, importance=T, ntree=50)
data.rf
```

A prediction is then generated.

```{r, echo=FALSE}
data.pred = predict(data.rf, testset)
table(data.pred, testset$classe)
confusionMatrix(data.pred, testset$classe)
```

Based on this exercise, the prediction model has produced an accuracy of 99.63 with kappa value of 0.9953

For validation, evaluated against the testing data. 

```{r}
preds = predict(data.rf, testing)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(preds)
```